Mathematical Aspects of the Capital Market

### 2017-08-25 phd/Chaptercorrelation.tex
Probably the final revision of the thesis.

Asymptotic theory for the eigenvalues of $\X\X'$ in the case of an entry distribution with infinite fourth moment was studied in \cite{soshnikov:2004,soshnikov:2006,auffinger:arous:peche:2009} in the cases when $p/n\to \gamma \in (0,\infty)$, while the authors of \cite{davis:mikosch:heiny:xie:2015,heiny:mikosch:2015:iid} allowed nearly arbitrary growth of the dimension $p$. 
In their model, the entries of $\X$ are regularly varying with index $\alpha>0$, implying that
\begin{equation}\label{eq:reg}
\P(|X|>x) = x^{-\alpha} L(x)\,,
\end{equation}
for a slowly varying function $L$. For $\alpha \in (0,4)$,  which implies an infinite fourth moment, they showed that $(a_{np}^{-2} \la_{(1)})$ converges to a \Frechet distributed random variable $\eta_{\alpha/2}$ with parameter $\alpha/2$ while $a_{np}^{-2} \la_{(p)}\cip 0$. Here the normalizing sequence $(a_n)$  is defined via $\P(|X|>a_n)\sim n^{-1}$, hence $n/a_{np}^{2}\to 0$.

@@

In view of \cite{davis:pfaffel:stelzer:2014,davis:mikosch:pfaffel:2015,davis:mikosch:heiny:xie:2015} the behavior of the top eigenvalues is reasonably well understood in the case of linear dependence among the $X_{it}$ and $\E[X^4]=\infty$. 
If $\E[X^4]<\infty$, similar arguments to \eqref{eq:lamu} can be developed to show that methods for sample covariance matrices can be applied to sample correlation matrices; see for example \cite{elkaroui:2009}. Theorem 1 in \cite{elkaroui:2009} proves that if the spectral norm of the population correlation matrix is uniformly bounded and $\E[X^4 (\log X)^{2+\vep}]<\infty$, then the spectral properties of $\bfR$ and $n^{-1} \X\X'$ are asymptotically the same. In particular, if $\lambda_{(1)}/n\cas c$, then $\mu_{(1)}\cas c$.

@@

In Figure~\ref{fig:stability}, we choose $X$ from the standardized
$t_{2.1}$-\ds , moving closer to the infinite variance case. The histogram of $(\mu_{(i)})$ fits the \MP density very well and the extreme eigenvalues are located in a close proximity of the endpoints of the \MP support. The sample covariance case in (b) does not 
look particularly appealing due to the fact that there are a few relatively large eigenvalues. For example, $\lambda_{(1)}/n=35.3196$ while $(1+\sqrt{\gamma})^2$ is only $1.7325$. By \cite{auffinger:arous:peche:2009,heiny:mikosch:2015:iid,davis:mikosch:heiny:xie:2015}, the properly normalized $\lambda_{(1)}$ converges to a \Frechet distributed random variable. The correct normalization is roughly $n^{4/2.1}$ and hence it is expected that $\lambda_{(1)}/n$ is separated from the bulk, whose behavior ultimately determines the limiting spectral distribution, which is the \MP law with parameter $\gamma=0.1$. However, due to the separation between the top eigenvalues and the bulk, it is not obvious from a histogram with only $50$ classes that the \MP law provides a good fit to the spectral distribution in (b). This different behavior of sample correlations and covariances is an additional argument for the higher stability of results obtained from an analysis of the sample correlation matrix.  
\par

>>>>

Asymptotic theory for the eigenvalues of $\X\X'$ in the case of an entry distribution with infinite fourth moment was studied in \cite{soshnikov:2004,soshnikov:2006,auffinger:arous:peche:2009} in the cases when $p/n\to \gamma \in (0,\infty)$, while the authors of \cite{davis:mikosch:heiny:xie:2015,heiny:mikosch:2016} allowed nearly arbitrary growth of the dimension $p$. 
In their model, the entries of $\X$ are regularly varying with index $\alpha>0$, implying that
\begin{equation}\label{eq:reg}
\P(|X|>x) = x^{-\alpha} L(x)\,,
\end{equation}
for a slowly varying function $L$. For $\alpha \in (0,4)$,  which implies an infinite fourth moment, they showed that $(a_{np}^{-2} \la_{(1)})$ converges to a \Frechet distributed random variable $\eta_{\alpha/2}$ with parameter $\alpha/2$ while $a_{np}^{-2} \la_{(p)}\cip 0$. Here the normalizing sequence $(a_n)$  is defined via $\P(|X|>a_n)\sim n^{-1}$, hence $n/a_{np}^{2}\to 0$.

@@

In view of \cite{davis:pfaffel:stelzer:2014,davis:mikosch:pfaffel:2016,davis:mikosch:heiny:xie:2015} the behavior of the top eigenvalues is reasonably well understood in the case of linear dependence among the $X_{it}$ and $\E[X^4]=\infty$. 
If $\E[X^4]<\infty$, similar arguments to \eqref{eq:lamu} can be developed to show that methods for sample covariance matrices can be applied to sample correlation matrices; see for example \cite{elkaroui:2009}. Theorem 1 in \cite{elkaroui:2009} proves that if the spectral norm of the population correlation matrix is uniformly bounded and $\E[X^4 (\log X)^{2+\vep}]<\infty$, then the spectral properties of $\bfR$ and $n^{-1} \X\X'$ are asymptotically the same. In particular, if $\lambda_{(1)}/n\cas c$, then $\mu_{(1)}\cas c$.

@@

In Figure~\ref{fig:stability}, we choose $X$ from the standardized
$t_{2.1}$-\ds , moving closer to the infinite variance case. The histogram of $(\mu_{(i)})$ fits the \MP density very well and the extreme eigenvalues are located in a close proximity of the endpoints of the \MP support. The sample covariance case in (b) does not 
look particularly appealing due to the fact that there are a few relatively large eigenvalues. For example, $\lambda_{(1)}/n=35.3196$ while $(1+\sqrt{\gamma})^2$ is only $1.7325$. By \cite{auffinger:arous:peche:2009,heiny:mikosch:2016,davis:mikosch:heiny:xie:2015}, the properly normalized $\lambda_{(1)}$ converges to a \Frechet distributed random variable. The correct normalization is roughly $n^{4/2.1}$ and hence it is expected that $\lambda_{(1)}/n$ is separated from the bulk, whose behavior ultimately determines the limiting spectral distribution, which is the \MP law with parameter $\gamma=0.1$. However, due to the separation between the top eigenvalues and the bulk, it is not obvious from a histogram with only $50$ classes that the \MP law provides a good fit to the spectral distribution in (b). This different behavior of sample correlations and covariances is an additional argument for the higher stability of results obtained from an analysis of the sample correlation matrix.  
\par

### 2017-08-25 phd/Chapteriidpaper.tex
Probably the final revision of the thesis.

Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} 

@@

\subsection{Structure of the paper}
The primary objective of this paper is to study the joint distribution of the largest eigenvalues of the sample covariance matrix $\z\z'$
in the case of iid regularly varying entries with infinite fourth moment.
We make a connection between extreme value theory, point process \con\ and the behavior of the largest eigenvalues. 
We study these eigenvalues under polynomial 
growth rates of the dimension $p$ relative to the sample size $n$. 
It turns out that they are essentially determined by the extreme diagonal elements of $\bfZ\bfZ'$ or, alternatively, 
by the extreme order statistics of the squared entries of $\bfZ$.  
\par
In Section~\ref{sec:2} we consider power-law growth rates of $(p_n)$, 
thereby generalizing proportional growth as prescribed by \eqref{eq:gamma}.
Our main results are presented in Section~\ref{sec:mainresults}. 
Theorem~\ref{thm:iidmain} provides approximations of the ordered eigenvalues of the sample covariance matrix either by
the ordered diagonal elements of $\bfZ\bfZ'$ or $\bfZ'\bfZ$, or by the order statistics of the squared entries of $\bfZ$.
These approximations provide a clear picture where the largest eigenvalues of the sample covariance matrix originate from. 
Our results generalize those in Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and 
Auffinger et al.~\cite{auffinger:arous:peche:2009} who assume proportionality of $p$ and $n$.
The employed techniques originate from extreme value analysis and large deviation theory; the proofs differ from those 
in the aforementioned literature. 
The same techniques can be applied when the entries of $\bfZ$ 
are heavy-tailed and allow for dependence through the rows and across the columns; see 
Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} for some recent attempts 
when the entries satisfy some linear dependence conditions. In the iid case, these results are covered by the present paper
and we also show that they remain valid under much more general growth conditions than in \cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}. In particular, we make clear that centering of the sample covariance matrix (as assumed in 
\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} when $Z$ has a finite second moment) is not needed. Thus, our techniques are applicable under rather general dependence structures. We refer to the recent work by Janssen et al.~\cite{janssen:mikosch:rezapour:xie:2016} on eigenvalues of stochastic volatility matrix models, where non-linear dependence was allowed.
% provided the rate of $p$ is sufficiently fast.

The convergence of the point processes of the properly normalized eigenvalues in Section~\ref{sec:ppconvergence} yields a multitude of useful findings connected to the joint distribution of the eigenvalues. As an application, the structure of the eigenvectors of $\z\z'$ is explored in Section~\ref{sec:eigenvectors}.
Technical proofs are collected in Section~\ref{sec:proofs}.
Section~\ref{sec:autocov} is devoted to an extension of the results to the singular values of the 
{\em sample autocovariance matrices}  which are a generalization of the traditional autocovariance function for time series to high-dimensional matrices. In applications, the analysis of sample autocovariance matrices for different lags might help to detect dependencies in the data; see Lam and Yao \cite{lam:yao} for related work.
 We conclude with Appendix~\ref{appendix:A} which contains useful facts about regular variation and point processes.

%------------------------------------------------------------------------
\section{Preliminaries}\label{sec:2}
In this section we will discuss growth rates for $p=p_n\to\infty$ and introduce some notation. 
\subsection{Growth rates for $p$}
In many applications it is not realistic to assume 
that the dimension $p$ of the data and the sample size $n$ grow at the same rate, i.e., condition \eqref{eq:gamma} is unlikely to be
satisfied.
The aforementioned results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al.~\cite{auffinger:arous:peche:2009} already show that the value $\gamma$ in the growth
rate~\eqref{eq:gamma} does not appear in the \ds al limits. 
This observation is in contrast to the light-tailed case; see \eqref{eq:MP} 
and \eqref{eq:geman}.  Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} allowed for more general rates for
$p_n\to\infty$ than linear growth in $n$. {\em However, they could not completely solve the technical difficulties arising with general growth rates of~$p$.}
%Recall that $p=p_n\to\infty$ is the number of rows in the
%matrix $\z_n$. 
In what follows, we 
specify the growth rate of $(p_n)$:
%a non-degenerate limit distribution of the normalized eigenvalues of the sample autocovariance matrices. 
%To be precise, we assume
\begin{equation}\label{eq:p}
p=p_n=n^\beta \ell(n), \qquad n\ge1,\tag{$C_p(\beta)$}
\end{equation}
where $\ell$ is a slowly varying function and $\beta\ge 0$. If $\beta =0$, we also assume $\ell(n) \to \infty$. 
Condition~\ref{eq:p} is more general than the growth conditions in the literature; see 
\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.

@@

In the right graph, we take the maxima as in \eqref{eq:rowa} and \eqref{eq:eigena} and show smoothed histograms of the approximation errors
$a_{np}^{-2}\max_{i\le p} |\lambda_{(i)}-D_{(i)}^\rightarrow|$, $a_{np}^{-2}\max_{i\le p}|\lambda_{(i)}-Z_{(i),np}^2|$. We take absolute values to deal with negative differences. 
Figure~\ref{fig:lambda_comparison} indicates 
that $(D_{(i)}^\rightarrow)$ yield a much better approximation to $(\la_{(i)})$ than $(Z_{(i),np}^2)$. Notice the different scaling on the $x$- and $y$-axes.
\begin{figure}[htb!]
  \centering
  \subfigure{
    \includegraphics[scale=0.34]{comparison_1.pdf}
  }
  \subfigure{
    \includegraphics[scale=0.34]{comparisonmax.pdf}
  }
  \caption{Smoothed histograms of the  approximation errors for the normalized eigenvalues $(a_{np}^{-2}\la_{(i)})$
for entries $Z_{it}$ with density \eqref{eq:distrsim}, $\alpha=1.6$, $\beta=1$, $n=1,000$ and $p=200$.
}
  \label{fig:lambda_comparison}
\end{figure}
\end{remark}
\par 
The proof of Theorem~\ref{thm:iidmain} will be given in Section~\ref{sec:proofs}. A main step in the proof is provided by the following
result whose proof will also be given in Section~\ref{sec:proofs};
a version of this theorem was proved in Davis et al.~\cite{davis:mikosch:pfaffel:2015} 
under more restrictive conditions on the growth rate of $(p_n)$.

@@

\subsection{Point process \con }\label{sec:ppconvergence}
In this section we want to illustrate how the approximations from Theorem~\ref{thm:iidmain} can be used
to derive \asy\ theory for the largest eigenvalues of $\z\z'$ via the weak \con\ of suitable \pp es.
The limiting \pp\ involves the points of the Poisson process
\begin{equation}\label{eq:N}
N_\Gamma=\sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\end{equation}
where $\vep_y$ is the Dirac measure at $y$, 
\beao%\label{eq:Gamma}
\Gamma_i=E_1+\cdots + E_i\,,\qquad i\ge 1\,,
\eeao
and $(E_i)$ is a sequence of iid standard exponential random variables. In other words, $N_\Gamma$ is a Poisson point process on $(0,\infty)$ with mean \ms\
$\mu(x,\infty)= x^{-\alpha/2}$, $x>0$.
\par
\begin{lemma}\label{lem:pp}
Assume the conditions of Theorem~\ref{prop:offdiagonal} hold. 
\begin{enumerate}
\item
If $\beta\ge 0$, then
\beam\label{eq:immb}
\sum_{i=1}^p \vep_{a_{np}^{-2}(D_i^\rightarrow-c_n)}\cid N_\Gamma\,,\qquad \nto\,,
\eeam  
where $c_n=0$ if $\E[D^\rightarrow]=\infty$ and $c_n=\E[D^\rightarrow]=n\,\E[Z^2]$ otherwise. 
\item
If $\beta\ge 0$, then
\beam\label{eq:imma}
\sum_{i=1}^p \vep_{a_{np}^{-2}Z_{(i),np}^2}\cid N_\Gamma\,,\qquad \nto\,,
\eeam  
\end{enumerate}
The weak \con\ of the \pp es holds in the space of 
point \ms s with state space $(0,\infty)$ equipped with the vague topology; see Resnick \cite{resnick:2007}.
\end{lemma}
\begin{remark} \rm
Similar results were used in the proofs of Davis et al.~\cite{davis:mikosch:heiny:xie:2015,davis:mikosch:pfaffel:2015}.
We also mention that the centering $c_n$ in the finite variance case can be avoided if $n/a_{np}^2\to 0$. 
The latter condition is satisfied if $\beta>\alpha/2-1$.
\end{remark}
\begin{proof} 
Part (1) follows from Lemma~\ref{lem:ppr}. As regards part (2), we observe that
\beam\label{eq:gp}
\sum_{i=1}^p\sum_{t=1}^n \vep_{a_{np}^{-2} Z_{it}^2}\std N_\Gamma\,;
\eeam
see e.g. Resnick~\cite{resnick:1987}, Proposition 3.21. On the other hand, $a_{np}^{-2} Z_{(p),np}^2\stp 0$ which together with
\eqref{eq:gp} yields part (2).
\end{proof}

Theorem~\ref{thm:iidmain}  and arguments similar to the proofs in Davis et al.~\cite{davis:mikosch:heiny:xie:2015,davis:mikosch:pfaffel:2015}
enable one to derive the weak \con\ of
the point processes of the normalized eigenvalues. 
%\begin{equation}\label{eq:ppdef}
%N_n=\sum_{i=1}^p \vep_{a_{np}^{-2} \lambda_i}, \quad \widetilde{N}_n^{\rightarrow}=\sum_{i=1}^p \vep_{a_{np}^{-2} (\lambda_i-n\,\E[Z^2])}, \qquad \mbox{and} \qquad
%\widetilde{N}_n^{\downarrow}=\sum_{i=1}^n \vep_{a_{np}^{-2} (\lambda_i-p\,\E[Z^2])}, \qquad n\ge 1\,.
%\end{equation}
\begin{theorem}\label{cor:1}
Assume the conditions of Theorem~\ref{thm:iidmain}. If
$\min(\beta,\beta^{-1} ) \in ((\alpha/2-1)_+,1]$ 
then
\beam\label{eq:imm}
\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i}\std N_\Gamma\,,
\eeam 
in the space of point measures 
with state space $(0,\infty)$ equipped with the vague topology. 
\end{theorem}
\begin{proof}
The limit relation \eqref{eq:imm} follows from \eqref{eq:imma} in combination with
\eqref{eq:eigena}. Alternatively, one can exploit \eqref{eq:immb} both for $(D_i^\rightarrow)$ and 
$(D_t^\downarrow)$ (notice that the \pp\ \con\ for the latter \seq\ follows by interchanging the roles of $n$ and $p$), the fact that
$(n\vee p)/a_{np}^2\to 0$ if $\min(\beta,\beta^{-1})\in ((\alpha/2-1)_+,1]$ (hence centering of the points $(D_i^\rightarrow)$ and  $(D_t^\downarrow)$ in \eqref{eq:immb}
can be avoided for $\E[Z^2]<\infty$) and finally using the approximations \eqref{eq:rowa} or \eqref{eq:cola}.
\end{proof}
The weak \con\ of the \pp es of the normalized eigenvalues of $\bfZ\bfZ'$ 
in Theorem~\ref{cor:1} allows one to use the conventional tools in this
field; see Resnick~\cite{resnick:2007,resnick:1987}. An immediate con\seq\ is
\beam \label{eq:helpgamma1}
a_{np}^{-2}\big(\la_{(1)},\ldots,\la_{(k)}\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)
\eeam
for any fixed $k\ge 1$. Using the methods of Davis et al.~\cite{davis:mikosch:heiny:xie:2015} shows for $\alpha \in (2,4)$
\beam \label{eq:helpgamma2}
a_{np}^{-2}\big(\la_{(1)}-(p\vee n) \E[Z^2],\ldots,\la_{(k)}-(p\vee n) \E[Z^2]\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,.
\eeam
Equations \eqref{eq:helpgamma1} and \eqref{eq:helpgamma2} yield that for $\alpha \in (0,4)$ and any fixed $k\ge 1$,
\beam \label{eq:helpgamma3}
a_{np}^{-2}\big(\la_{(1)}- \la_{(2)},\ldots, \la_{(k)}- \la_{(k+1)}\big)\std \big(\Gamma_1^{-2/\alpha} - \Gamma_{2}^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha} - \Gamma_{k+1}^{-2/\alpha} \big)\,.
\eeam
Related results can also be derived for an increasing number of order statistics, e.g.~the joint \con\ of the largest eigenvalue $a_{np}^{-2}\la_{(1)}$ and the trace $a_{np}^{-2}(\la_1+\cdots+\la_p)$. In particular, one obtains for
$\alpha\in (0,2)$ under the conditions of Theorem~\ref{cor:1} that
\beao
\dfrac{\la_{(1)}}{\la_1+\cdots +\la_p} \std \dfrac{\Gamma_1^{-2/\alpha}}{\Gamma_1^{-2/\alpha}+\Gamma_2^{-2/\alpha}+\cdots}\,.
\eeao 
We refer to Davis et al.~\cite{davis:mikosch:pfaffel:2015} for details on the proofs and more examples.

>>>>

Davis et al.~\cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014} 

@@

\subsection{Structure of the paper}
The primary objective of this paper is to study the joint distribution of the largest eigenvalues of the sample covariance matrix $\z\z'$
in the case of iid regularly varying entries with infinite fourth moment.
We make a connection between extreme value theory, point process \con\ and the behavior of the largest eigenvalues. 
We study these eigenvalues under polynomial 
growth rates of the dimension $p$ relative to the sample size $n$. 
It turns out that they are essentially determined by the extreme diagonal elements of $\bfZ\bfZ'$ or, alternatively, 
by the extreme order statistics of the squared entries of $\bfZ$.  
\par
In Section~\ref{sec:2} we consider power-law growth rates of $(p_n)$, 
thereby generalizing proportional growth as prescribed by \eqref{eq:gamma}.
Our main results are presented in Section~\ref{sec:mainresults}. 
Theorem~\ref{thm:iidmain} provides approximations of the ordered eigenvalues of the sample covariance matrix either by
the ordered diagonal elements of $\bfZ\bfZ'$ or $\bfZ'\bfZ$, or by the order statistics of the squared entries of $\bfZ$.
These approximations provide a clear picture where the largest eigenvalues of the sample covariance matrix originate from. 
Our results generalize those in Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and 
Auffinger et al.~\cite{auffinger:arous:peche:2009} who assume proportionality of $p$ and $n$.
The employed techniques originate from extreme value analysis and large deviation theory; the proofs differ from those 
in the aforementioned literature. 
The same techniques can be applied when the entries of $\bfZ$ 
are heavy-tailed and allow for dependence through the rows and across the columns; see 
Davis et al.~\cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014} for some recent attempts 
when the entries satisfy some linear dependence conditions. In the iid case, these results are covered by the present paper
and we also show that they remain valid under much more general growth conditions than in \cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014}. In particular, we make clear that centering of the sample covariance matrix (as assumed in 
\cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014} when $Z$ has a finite second moment) is not needed. Thus, our techniques are applicable under rather general dependence structures. We refer to the recent work by Janssen et al.~\cite{janssen:mikosch:rezapour:xie:2016} on eigenvalues of stochastic volatility matrix models, where non-linear dependence was allowed.
% provided the rate of $p$ is sufficiently fast.

The convergence of the point processes of the properly normalized eigenvalues in Section~\ref{sec:ppconvergence} yields a multitude of useful findings connected to the joint distribution of the eigenvalues. As an application, the structure of the eigenvectors of $\z\z'$ is explored in Section~\ref{sec:eigenvectors}.
Technical proofs are collected in Section~\ref{sec:proofs}.
Section~\ref{sec:autocov} is devoted to an extension of the results to the singular values of the 
{\em sample autocovariance matrices}  which are a generalization of the traditional autocovariance function for time series to high-dimensional matrices. In applications, the analysis of sample autocovariance matrices for different lags might help to detect dependencies in the data; see Lam and Yao \cite{lam:yao:2012} for related work.
 We conclude with Appendix~\ref{appendix:A} which contains useful facts about regular variation and point processes.

%------------------------------------------------------------------------
\section{Preliminaries}\label{sec:2}
In this section we will discuss growth rates for $p=p_n\to\infty$ and introduce some notation. 
\subsection{Growth rates for $p$}
In many applications it is not realistic to assume 
that the dimension $p$ of the data and the sample size $n$ grow at the same rate, i.e., condition \eqref{eq:gamma} is unlikely to be
satisfied.
The aforementioned results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al.~\cite{auffinger:arous:peche:2009} already show that the value $\gamma$ in the growth
rate~\eqref{eq:gamma} does not appear in the \ds al limits. 
This observation is in contrast to the light-tailed case; see \eqref{eq:MP} 
and \eqref{eq:geman}.  Davis et al.~\cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014} allowed for more general rates for
$p_n\to\infty$ than linear growth in $n$. {\em However, they could not completely solve the technical difficulties arising with general growth rates of~$p$.}
%Recall that $p=p_n\to\infty$ is the number of rows in the
%matrix $\z_n$. 
In what follows, we 
specify the growth rate of $(p_n)$:
%a non-degenerate limit distribution of the normalized eigenvalues of the sample autocovariance matrices. 
%To be precise, we assume
\begin{equation}\label{eq:p}
p=p_n=n^\beta \ell(n), \qquad n\ge1,\tag{$C_p(\beta)$}
\end{equation}
where $\ell$ is a slowly varying function and $\beta\ge 0$. If $\beta =0$, we also assume $\ell(n) \to \infty$. 
Condition~\ref{eq:p} is more general than the growth conditions in the literature; see 
\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014}.

@@

In the right graph, we take the maxima as in \eqref{eq:rowa} and \eqref{eq:eigena} and show smoothed histograms of the approximation errors
$a_{np}^{-2}\max_{i\le p} |\lambda_{(i)}-D_{(i)}^\rightarrow|$, $a_{np}^{-2}\max_{i\le p}|\lambda_{(i)}-Z_{(i),np}^2|$. We take absolute values to deal with negative differences. 
Figure~\ref{fig:lambda_comparison} indicates 
that $(D_{(i)}^\rightarrow)$ yield a much better approximation to $(\la_{(i)})$ than $(Z_{(i),np}^2)$. Notice the different scaling on the $x$- and $y$-axes.
\begin{figure}[htb!]
  \centering
  \subfigure{
    \includegraphics[scale=0.34]{comparison_1.pdf}
  }
  \subfigure{
    \includegraphics[scale=0.34]{comparisonmax.pdf}
  }
  \caption{Smoothed histograms of the  approximation errors for the normalized eigenvalues $(a_{np}^{-2}\la_{(i)})$
for entries $Z_{it}$ with density \eqref{eq:distrsim}, $\alpha=1.6$, $\beta=1$, $n=1,000$ and $p=200$.
}
  \label{fig:lambda_comparison}
\end{figure}
\end{remark}
\par 
The proof of Theorem~\ref{thm:iidmain} will be given in Section~\ref{sec:proofs}. A main step in the proof is provided by the following
result whose proof will also be given in Section~\ref{sec:proofs};
a version of this theorem was proved in Davis et al.~\cite{davis:mikosch:pfaffel:2016} 
under more restrictive conditions on the growth rate of $(p_n)$.

@@

\subsection{Point process \con }\label{sec:ppconvergence}
In this section we want to illustrate how the approximations from Theorem~\ref{thm:iidmain} can be used
to derive \asy\ theory for the largest eigenvalues of $\z\z'$ via the weak \con\ of suitable \pp es.
The limiting \pp\ involves the points of the Poisson process
\begin{equation}\label{eq:N}
N_\Gamma=\sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\end{equation}
where $\vep_y$ is the Dirac measure at $y$, 
\beao%\label{eq:Gamma}
\Gamma_i=E_1+\cdots + E_i\,,\qquad i\ge 1\,,
\eeao
and $(E_i)$ is a sequence of iid standard exponential random variables. In other words, $N_\Gamma$ is a Poisson point process on $(0,\infty)$ with mean \ms\
$\mu(x,\infty)= x^{-\alpha/2}$, $x>0$.
\par
\begin{lemma}\label{lem:pp}
Assume the conditions of Theorem~\ref{prop:offdiagonal} hold. 
\begin{enumerate}
\item
If $\beta\ge 0$, then
\beam\label{eq:immb}
\sum_{i=1}^p \vep_{a_{np}^{-2}(D_i^\rightarrow-c_n)}\cid N_\Gamma\,,\qquad \nto\,,
\eeam  
where $c_n=0$ if $\E[D^\rightarrow]=\infty$ and $c_n=\E[D^\rightarrow]=n\,\E[Z^2]$ otherwise. 
\item
If $\beta\ge 0$, then
\beam\label{eq:imma}
\sum_{i=1}^p \vep_{a_{np}^{-2}Z_{(i),np}^2}\cid N_\Gamma\,,\qquad \nto\,,
\eeam  
\end{enumerate}
The weak \con\ of the \pp es holds in the space of 
point \ms s with state space $(0,\infty)$ equipped with the vague topology; see Resnick \cite{resnick:2007}.
\end{lemma}
\begin{remark} \rm
Similar results were used in the proofs of Davis et al.~\cite{davis:mikosch:heiny:xie:2015,davis:mikosch:pfaffel:2016}.
We also mention that the centering $c_n$ in the finite variance case can be avoided if $n/a_{np}^2\to 0$. 
The latter condition is satisfied if $\beta>\alpha/2-1$.
\end{remark}
\begin{proof} 
Part (1) follows from Lemma~\ref{lem:ppr}. As regards part (2), we observe that
\beam\label{eq:gp}
\sum_{i=1}^p\sum_{t=1}^n \vep_{a_{np}^{-2} Z_{it}^2}\std N_\Gamma\,;
\eeam
see e.g. Resnick~\cite{resnick:1987}, Proposition 3.21. On the other hand, $a_{np}^{-2} Z_{(p),np}^2\stp 0$ which together with
\eqref{eq:gp} yields part (2).
\end{proof}

Theorem~\ref{thm:iidmain}  and arguments similar to the proofs in Davis et al.~\cite{davis:mikosch:heiny:xie:2015,davis:mikosch:pfaffel:2016}
enable one to derive the weak \con\ of
the point processes of the normalized eigenvalues. 
%\begin{equation}\label{eq:ppdef}
%N_n=\sum_{i=1}^p \vep_{a_{np}^{-2} \lambda_i}, \quad \widetilde{N}_n^{\rightarrow}=\sum_{i=1}^p \vep_{a_{np}^{-2} (\lambda_i-n\,\E[Z^2])}, \qquad \mbox{and} \qquad
%\widetilde{N}_n^{\downarrow}=\sum_{i=1}^n \vep_{a_{np}^{-2} (\lambda_i-p\,\E[Z^2])}, \qquad n\ge 1\,.
%\end{equation}
\begin{theorem}\label{cor:1}
Assume the conditions of Theorem~\ref{thm:iidmain}. If
$\min(\beta,\beta^{-1} ) \in ((\alpha/2-1)_+,1]$ 
then
\beam\label{eq:imm}
\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i}\std N_\Gamma\,,
\eeam 
in the space of point measures 
with state space $(0,\infty)$ equipped with the vague topology. 
\end{theorem}
\begin{proof}
The limit relation \eqref{eq:imm} follows from \eqref{eq:imma} in combination with
\eqref{eq:eigena}. Alternatively, one can exploit \eqref{eq:immb} both for $(D_i^\rightarrow)$ and 
$(D_t^\downarrow)$ (notice that the \pp\ \con\ for the latter \seq\ follows by interchanging the roles of $n$ and $p$), the fact that
$(n\vee p)/a_{np}^2\to 0$ if $\min(\beta,\beta^{-1})\in ((\alpha/2-1)_+,1]$ (hence centering of the points $(D_i^\rightarrow)$ and  $(D_t^\downarrow)$ in \eqref{eq:immb}
can be avoided for $\E[Z^2]<\infty$) and finally using the approximations \eqref{eq:rowa} or \eqref{eq:cola}.
\end{proof}
The weak \con\ of the \pp es of the normalized eigenvalues of $\bfZ\bfZ'$ 
in Theorem~\ref{cor:1} allows one to use the conventional tools in this
field; see Resnick~\cite{resnick:2007,resnick:1987}. An immediate con\seq\ is
\beam \label{eq:helpgamma1}
a_{np}^{-2}\big(\la_{(1)},\ldots,\la_{(k)}\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)
\eeam
for any fixed $k\ge 1$. Using the methods of Davis et al.~\cite{davis:mikosch:heiny:xie:2015} shows for $\alpha \in (2,4)$
\beam \label{eq:helpgamma2}
a_{np}^{-2}\big(\la_{(1)}-(p\vee n) \E[Z^2],\ldots,\la_{(k)}-(p\vee n) \E[Z^2]\big)\std \big(\Gamma_1^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha}\big)\,.
\eeam
Equations \eqref{eq:helpgamma1} and \eqref{eq:helpgamma2} yield that for $\alpha \in (0,4)$ and any fixed $k\ge 1$,
\beam \label{eq:helpgamma3}
a_{np}^{-2}\big(\la_{(1)}- \la_{(2)},\ldots, \la_{(k)}- \la_{(k+1)}\big)\std \big(\Gamma_1^{-2/\alpha} - \Gamma_{2}^{-2/\alpha},\ldots,\Gamma_k^{-2/\alpha} - \Gamma_{k+1}^{-2/\alpha} \big)\,.
\eeam
Related results can also be derived for an increasing number of order statistics, e.g.~the joint \con\ of the largest eigenvalue $a_{np}^{-2}\la_{(1)}$ and the trace $a_{np}^{-2}(\la_1+\cdots+\la_p)$. In particular, one obtains for
$\alpha\in (0,2)$ under the conditions of Theorem~\ref{cor:1} that
\beao
\dfrac{\la_{(1)}}{\la_1+\cdots +\la_p} \std \dfrac{\Gamma_1^{-2/\alpha}}{\Gamma_1^{-2/\alpha}+\Gamma_2^{-2/\alpha}+\cdots}\,.
\eeao 
We refer to Davis et al.~\cite{davis:mikosch:pfaffel:2016} for details on the proofs and more examples.
